---
title: "DeFi Example Notebook: AAVE Users"
subtitle: "AAVE Users Analysis"
author: "Data Analytics Research Instructors"
date: "08/27/2021"
output:
  pdf_document: default
  html_document:
    toc: true
    number_sections: true
    df_print: paged
---

```{r}
#import libraries
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
library(ggbiplot)
library(gplots)
library(RColorBrewer)
library(beeswarm)
library(tidyverse)
library(tidyquant) 
library(ggbeeswarm)
library(foreach)
library(doParallel)
library(Rtsne)
library(anytime)
```
We begin by loading the csv file in to a dataframe.
```{r}
#load in csv file to data frame
df<-read_rds("~/transactionsv2.rds")
head(df)
```
We reformat the dataframe so that each row represents a user, and the columns represent a summarization of the user's transaction history. 
```{r}
#group by user and get time of user's first and last transaction, as well as number of transactions
df.users<- df%>%group_by(user, user_alias)%>%
  summarise(timefirst=min(anydate(timestamp)), timelast=max(anydate(timestamp)), N=n())
#get the time the user has been active
df.users$timeactive<-df.users$timelast-df.users$timefirst
#get amounts for columns
df$logUSD<-log10(df$amountUSD)
df$logCollateralUSD<-log10(df$amountUSDCollateral)
#get user's transaction information
for(Type in unique(df$type)){
  #filter for only transactions of certain type
  df.type <-filter(df%>%group_by(user)%>%
                     count(type),type==Type)
  
  #add means of each transaction type
  if(Type!="liquidation" || Type!="swap"){
    df.sum<-filter(df,type==Type)%>%
      group_by(user)%>%
      summarise(Sum=sum(logUSD))
    colnames(df.sum)[2]<-paste('total_',Type,sep='')
    df.users<-merge(x=df.users,y=df.sum,by="user",all.x=TRUE)
  } 
  
  #add counts of transaction types to df
  ntypes<-paste("n",Type,sep='')
  colnames(df.type)[3]<-ntypes
  df.users<-merge(x=df.users,y=select(df.type,user,ntypes),by="user",all.x=TRUE)
  
  #get proportion of transaction types and weekly number of transaction type
  df.users[paste("prop_",Type,sep='')]<-(df.users[ntypes]+.05)/((df.users$N)+.3)
}
head(df.users)
```
```{r}
nrow(df.users)
length(unique(df.users$user))
```


As long as those two numbers are the same, we are good to proceed. Next, we select only the columns we wish to cluster on.
```{r}
#subset only columns we wish to scale by removing columns that we will not cluster on
df.sub<-select(df.users,-c(user,timefirst,timelast,nborrow,nrepay,nswap,nliquidation,nredeem,ndeposit, total_swap, nswap, total_collateral, total_liquidation, ncollateral, user_alias))
#repalce missing values as 0's
df.sub<-df.sub%>%replace(is.na(.),0)
```
We scale the data to prepare for PCA.
```{r}
#scale data
df.scaled<-df.sub%>%mutate_all(scale)

summary(df.scaled)
```
Now, we perform PCA on the scaled data. 
```{r}
df.scaled <- df.scaled[-c(41100, 44263, 44658, 14211, 20033, 6638),]
df.users <- df.users[-c(41100, 44263, 44658, 14211, 20033, 6638),]

#perform pca on data
my.pca<-prcomp(df.scaled,retx=TRUE,center=FALSE,scale=FALSE) # Run PCA and save to my.pca
#make scree plot
plot(my.pca, type="line")
#summary of pca
summary(my.pca)
ncomps=5
```

Finally, we run the kmeans clustering algorithm on the data. We select the number of clusters equal to the number of selected components. 
```{r}
#run kmeans algorithm
set.seed(1)
km <-kmeans(df.scaled,4)
#plot frequencies of each cluster
barplot(table(km$cluster),main="Kmeans Cluster Size")
km$size
```
Finally, we view the centers of the kmeans clusters.
```{r}
#make heatmap of cluster centers
heatmap.2(km$centers,
scale = "col",
dendrogram = "row",
Colv=FALSE,
cexCol=1.0,
main = "Kmeans Cluster Centers",
trace ="none")
```
# Exploring the Influential Factors with a Biplot 
```{r}
plot1<-ggbiplot(my.pca,choices=c(1,2),
  labels=rownames(df.scaled), #show point labels
  var.axes=TRUE, # Display axes
  ellipse = FALSE, # Don't display ellipse
  obs.scale=1,
  groups=as.factor(km$cluster)) +
  ggtitle("User Data Projected on PC1 and PC2 ")
plot1
```


```{r}
df["date"] = anydate(df$timestamp)
min(df$date)
max(df$date)
num_days <- max(df$date) - min(df$date)
num_days
```

```{r}
ggplot(data = df, aes(x = date,  group = type, color = type)) + 
  geom_density()+
  geom_vline(xintercept = as.numeric(as.Date("2021-05-18")), linetype=2, alpha = 0.5)+
  ggtitle("Transaction Types Over Time")
```





```{r}
df.clusters <- data.frame(user = df.users$user, cluster = km$cluster)
df <- left_join(df, df.clusters, by = "user")
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
# 6-list of ggplot colors explicitly specified
pgg <- gg_color_hue(6)
```

```{r}
cluster_names <- c('1' = "Cluster 1", '2' = "Cluster 2", '3' = "Cluster 3", '4' = "Cluster 4")
ggplot(data = df[!(is.na(df$cluster)) & (df$type != "collateral"), ], aes(x = date,  group = type, color = type)) + 
  geom_density()+
  ggtitle("Transaction Types Over Time by User Cluster")+
  geom_vline(xintercept = as.numeric(as.Date("2021-05-18")), linetype=2, alpha = 0.5, color = "black")+
  scale_x_date(date_breaks = "2 months", date_labels = "%b-%y")+
  scale_color_manual("type", values = c("deposit"="green","borrow" = pgg[4], "redeem" = "yellow", "liquidation" = "red","repay"=pgg[6],"swap"=pgg[5]))+
  scale_fill_manual("type", values = c("deposit"="green","borrow" = pgg[4], "redeem" = "yellow", "liquidation" = "red","repay"=pgg[6],"swap"=pgg[5]))+
  facet_wrap(~ cluster, labeller = as_labeller(cluster_names))
  
```

```{r}
df <- filter(df, cluster != "NA")
ggplot(df, aes(x = date)) + 
  geom_histogram(binwidth = 7, color = "black", fill = "white")+
  geom_vline(xintercept = as.numeric(as.Date("2021-05-18")), linetype=2, alpha = 0.8, color = "red")+
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y")+
  ggtitle("Total Number of Transactions per Week")
```

```{r}
ggplot(filter(df, type == "liquidation"), aes(x = date)) + 
  geom_histogram(binwidth = 1, color = "black", fill = "white")+
  geom_vline(xintercept = as.numeric(as.Date("2021-05-18")), linetype=2, alpha = 0.8, color = "red")+
    scale_x_date(date_breaks = "1 month", date_labels = "%b-%y")+
  ggtitle("Total Number of Liquidations per Day")
```

```{r}
ggplot(df, aes(x = date)) + 
  geom_histogram(binwidth = 1, color = "black", fill = "white")+
  geom_vline(xintercept = as.numeric(as.Date("2021-05-18")), linetype=2, alpha = 0.8, color = "red")+
  ggtitle("Total Number of Transactions per Day")+
  facet_wrap(~ type)
```

```{r}
ggplot(filter(df, type == "swap"), aes(x = date)) + 
  geom_histogram(binwidth = 1, color = "black", fill = "white")+
  geom_vline(xintercept = as.numeric(as.Date("2021-05-18")), linetype=2, alpha = 0.8, color = "red")+
  ggtitle("Total Number of Swaps per Day")
```

```{r}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
```

```{r}

p <- ggplot(filter(df, type == "deposit"), aes(x = date, 
                                            text = paste('<Br>Date: ', as.Date(date)))) + 
  geom_histogram(binwidth = 1, color = "black", fill = "white")+
  geom_vline(xintercept = as.numeric(as.Date("2021-05-18")), linetype=2, alpha = 0.8, color = "red")+
  ggtitle("Total Number of Deposits per Day") 
ggplotly(p, tooltip = c("y", "text")) 
```



```{r}
p <- ggplot(filter(df, type == "deposit"), aes(x = date,
                                            text = paste('<Br>Date: ', as.Date(date)))) + 
  geom_histogram(binwidth = 7, color = "black", fill = "white")+
  geom_vline(xintercept = as.numeric(as.Date("2021-05-18")), linetype=2, alpha = 0.8, color = "red")+
  ggtitle("Total Number of Deposits per Week")+
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y")
ggplotly(p, tooltip = c("y", "text")) 
```

```{r}
ggplot(data = filter(df, type == "deposit"), aes(x = date, y = amountUSD, fill = as.factor(cluster)))+
  geom_col()+
  geom_vline(xintercept = as.numeric(as.Date("2021-05-18")), linetype=2, alpha = 0.8, color = "red")+
  ggtitle("Amount of Money (USD) Deposited per Day")+
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y")

```

```{r}
ggplot(filter(df, type == "deposit"), aes(x = date)) + 
  geom_histogram(binwidth = 7, color = "black", fill = "white")+
  stat_bin(binwidth=7, geom='text', color='black', aes(label=..count..),
           position=position_stack(vjust = 1.035))+
  geom_vline(xintercept = as.numeric(as.Date("2021-05-18")), linetype=2, alpha = 0.8, color = "red")+
  ggtitle("Total Number of Deposits per Week")+
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y")



```

```{r}
threshold <- 50
user_freq <- df %>% 
  count(onBehalfOf_alias)
user_freq <- filter(user_freq, n >= threshold)
top_users <- user_freq$onBehalfOf_alias
```

```{r}
example_user <- sample(top_users, 1)
user_transactions <- filter(df, onBehalfOf_alias == example_user)
ggplot(user_transactions, aes(x = date)) + 
  geom_histogram(binwidth = 1, color = "black", fill = "white")+
  geom_vline(xintercept = as.numeric(as.Date("2021-05-18")), linetype=2, alpha = 0.8, color = "red")+
  ggtitle(example_user, "Total Number of Transactions per Day")+
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y")+
  facet_wrap(~ type)
```

```{r}
first <- min(as.Date(user_transactions$date))
recent <- max(user_transactions$date)
num <- nrow(user_transactions)
most_active <- names(sort(-table(user_transactions$date)))[1]
cat(paste(example_user))
cat(paste("\nFirst Transaction: ", first))
cat(paste("\nMost Recent Transaction: ", recent))
cat(paste("\nNumber of Transactions: ", num))
cat(paste("\nMost Active Day: ", most_active))
types_counts <- user_transactions %>%
  count(type) %>%
  mutate(n / (as.numeric(recent - first)/7))
types_counts <- rename(types_counts, "Total" = "n")
types_counts <- rename(types_counts, "Weekly Average" = "n/(as.numeric(recent - first)/7)")
kable(types_counts)
```

```{r}
df.weekly<- df%>%group_by(user, user_alias)%>%
  summarise(timefirst=min(anydate(timestamp)), timelast=max(anydate(timestamp)), N=n())
#get the time the user has been active
df.weekly$timeactive<-df.weekly$timelast-df.weekly$timefirst
#get amounts for columns
df$logUSD<-log10(df$amountUSD)
df$logCollateralUSD<-log10(df$amountUSDCollateral)
#get user's transaction information
for(Type in unique(df$type)){
  #filter for only transactions of certain type
  df.type <-filter(df%>%group_by(user)%>%
                     count(type),type==Type)
  
  #add means of each transaction type
  if(Type!="liquidation" || Type!="swap"){
    df.sum<-filter(df,type==Type)%>%
      group_by(user)%>%
      summarise(Sum=sum(logUSD))
    colnames(df.sum)[2]<-paste('total_',Type,sep='')
    df.weekly<-merge(x=df.weekly,y=df.sum,by="user",all.x=TRUE)
  } 
  
  #add counts of transaction types to df
  ntypes<-paste("n",Type,sep='')
  colnames(df.type)[3]<-ntypes
  df.weekly<-merge(x=df.weekly,y=select(df.type,user,ntypes),by="user",all.x=TRUE)
}
df.weekly <- rename(df.weekly, "name" = "user_alias")

head(df.weekly)
df.weekly$timeactive =as.numeric( df.weekly$timeactive / 7)
df.weekly <- rename(df.weekly, "weeks" = "timeactive")

```



```{r}
one_day <- df.weekly %>%
  filter(weeks == 0)
df.weekly <- df.weekly %>%
  filter(weeks > 0)
df.weekly$N =df.weekly$N / df.weekly$weeks
df.weekly[,7:20] <- df.weekly[,7:20] / df.weekly$weeks
one_day$N <- one_day$N * 7
one_day[,7:20] <- one_day[,7:20] * 7
df.weekly <- rbind(df.weekly, one_day)
df.weekly <- rename(df.weekly, "weekly_n" = "N")
head(df.weekly)
liquids <- df.weekly %>%
  filter(nliquidation > 0)
```

```{r}
df.sub<-select(df.weekly,-c(user,name, total_liquidation, total_swap, total_collateral))
#repalce missing values as 0's
df.sub[is.na(df.sub)] <- 0
#scale data
df.scaled<-df.sub%>%mutate_all(scale)

summary(df.scaled)
```
Now, we perform PCA on the scaled data. 
```{r}
#perform pca on data
my.pca<-prcomp(df.scaled,retx=TRUE,center=FALSE,scale=FALSE) # Run PCA and save to my.pca
#make scree plot
plot(my.pca, type="line")
#summary of pca
summary(my.pca)
ncomps=5
```


```{r}
#run kmeans algorithm
set.seed(1)
km <-kmeans(df.scaled,7)
#plot frequencies of each cluster
barplot(table(km$cluster),main="Kmeans Cluster Size")
km$size
```
Finally, we view the centers of the kmeans clusters.
```{r}
#make heatmap of cluster centers
heatmap.2(km$centers,
scale = "col",
dendrogram = "row",
Colv=FALSE,
cexCol=1.0,
main = "Kmeans Cluster Centers",
trace ="none")
```
# Exploring the Influential Factors with a Biplot 
```{r}
plot1<-ggbiplot(my.pca,choices=c(1,2),
  #labels=rownames(df.scaled), #show point labels
  var.axes=TRUE, # Display axes
  ellipse = FALSE, # Don't display ellipse
  obs.scale=1,
  groups=as.factor(km$cluster)) +
  ggtitle("User Data Projected on PC1 and PC2 ")
plot1
```



```{r}
ggplot(df.weekly %>% filter(weekly_n <= 100), aes(x = weekly_n))+
  geom_histogram(binwidth = 5, color = "black", fill = "white", aes(y = ..count../sum(..count..)))+
  ggtitle("Number of Transactions per Week for Users with less than 100")+
  scale_y_continuous(labels=percent)+
  labs(y = "Percent of Users", x = "Number of Transactions")
```

```{r}
df.clusters <- data.frame(cluster = km$cluster, name = df.weekly$name)
df.clusters <- left_join(df.weekly, df.clusters, by = "name")
```

```{r}
mean(df.weekly$weekly_n)
for (x in 1:7){
  cluster_x <- df.clusters %>%
    filter(cluster == x)
  print(paste("Cluster",x, ":",  round(mean(cluster_x$weekly_n), 2)))
}
```

