---
title: "DAR Defi Team Assignment 2 (Fall 2021)"
subtitle: "DeFi Reserve Coins"
author: "Aaron Micah Green"
date: "09/08/2021"
output:
  pdf_document: default
  html_document:
    toc: true
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
# Set the default CRAN repository
local({r <- getOption("repos")
       r["CRAN"] <- "http://cran.r-project.org" 
       options(repos=r)
})

# Set code chunk defaults
knitr::opts_chunk$set(echo = TRUE)

# Load required packages; install if necessary
# CAUTION: DO NOT interrupt R as it installs packages!!
if (!require("ggplot2")) {
  install.packages("ggplot2")
  library(ggplot2)
}

if (!require("knitr")) {
  install.packages("knitr")
  library(knitr)
}

if (!require("dplyr")) {
  install.packages("dplyr")
  library(dp)
}

if (!require("RColorBrewer")) {
  install.packages("RColorBrewer")
  library(RColorBrewer)
}
if (!require("beeswarm")) {
  install.packages("beeswarm")
  library(beeswarm)
}
if (!require("tidyverse")) {
  install.packages("tidyverse")
  library(tidyverse)
}
if (!require("ggbeeswarm")) {
  install.packages("ggbeeswarm")
  library(ggbeeswarm)
}
if (!require("xts")) {
  install.packages("xts")
  library(xts)
}

```
# Prepare Transaction Data and Explore

We begin by loading our prepared AAVE transaction data into a dataframe. The dataset has over 400,000 rows, and 27 columns. 

We are directly loading the dataframe from an Rds archive instead of a CSV file to conserve space. 

```{r}
#load Rds (binary version of csv file) into dataframe
# Assumes this notebook is in: ~/IDEA-Blockchain/DefiResearch/StudentNotebooks/Assignment02
df<-read_rds('../../Data/transactions.Rds')

# Let's take a quick look at the first few observation
head(df)
```

Now look at the summaries to see the types, values, and missingness (NA's) of the data.

```{r}

summary(df)
```


First we'll do some preliminary analysis before we ask detailed questions.

## Analyze Transaction Types

Let's examine the different types of _transactions_ present in the data. We'll make a simple bar plot to visualize the number of each transaction types. "Deposit" is the most common type of transaction, whereas "swaps" are the most rare.

```{r}
#set color palette
colors = brewer.pal(6,"Set2")

#create barplot
barplot(table(df$type), main='Transaction Type Counts', xlab='Type',ylab='Count',col=colors)
```

There are more "deposits" than "borrows," because users often need to overcollateralize for loans. 

Now we'll examine the amount of US dollars being used in the different types of transactions. We create box plots for the four types of transactions that have the "amount" feature associated with them, and we visualize the distribution of that column for the different transactions. 

We can see that most transactions are completed with very little money. 

```{r}
#create boxplot
boxplot(amountUSD~type,data=df,outline=FALSE,col=colors,
        main="Transaction Amounts",xlab="Type",ylab="Amount (USD)")

```

We do find some very large amounts, so it's helpful to look at this on a log scale.

```{r}
boxplot(log(amountUSD)~type,data=df,outline=FALSE,col=colors,
        main="Log Transaction Amounts",xlab="Type",ylab="Log Amount (USD)")
```

Observation: _There are many borrows and repays with high transactions amounts, but deposits and redeems have much lower transactions amounts._

## Examine Reserve Coins  

There are 50 different "Reserve" coins used in transactions in AAVE.
Let's create a table of those reserve coins with at least 500 transactions and rank order them by their volume. 

```{r}
# Use deplyr to drop NA reserves, add the counts and then kep only the top 20
reservecoins <- df %>%  drop_na(reserve) %>% 
  count(reserve) %>% 
  arrange(-n) %>% 
  head(20)

# Add the rank to help keep track of the reserve coins
reservecoins <- reservecoins %>% 
  mutate(rank=1:nrow(reservecoins),.before=reserve)

# List the results nicely with kable()
kable(reservecoins)

```

Let's look at the number of transactions types for each currency. 

```{r}

TopcoinSummary <- df %>% filter(reserve %in% reservecoins$reserve) %>%
  group_by(reserve) %>% 
  count(type) %>% 
  mutate(percent = n/sum(n)*100)

kable(TopcoinSummary) 

```


## Look at Sample User Transaction Histories 

Finally, we will examine the transaction history of different users. To do this, we will select 3 random users from the data who have completed between 100 and 300 transactions. Then, we create swarmplots displaying the different types of transactions those users made over time.

```{r,fig.width=16,fig.height=8}
#set seed
set.seed(1)

# Select three random users that have between 100 and 300 transactions
users<-vector(length=3)
count<-0
while(count<=3){
  success<-FALSE
  while(!success){
    #get random user
    ruser<-sample(df$user,1)
    
    #check for valid number of transactions
    length<-nrow(filter(df,user==ruser))
    if (length>100 && length<300){
      users[count]=ruser
      success<-TRUE
      count<-count+1
    }
  }
}
df.rusers<-filter(df, user %in%users)

# Create a "swarmplot"

ggplot(df.rusers,aes(user, timestamp,color=type)) + 
        geom_beeswarm(cex=1)+
        coord_flip()+
        ggtitle("User Transaction History")

```

Observation: _Users have very different transactions patterns, which we will try to better understand. _


# Activities
##$\color{blue}{Exercise~1} (10 pts)$  

1. Divide the top 20 reserve coins between your team members so each team member has the same amount of coins; feel free to add or subtract coins if necessary.
```{r}

```

2. Look up your coins on the internet to find out what they are.  ProTip: Look them up on http://defipulse.com to see their _Total Value Locked (TVL)_.

3. Examine the percentage of transaction types for your coins.  Hypothesize why a given coin might have more of one type of transaction than another.

4. _Prepare one slide with the findings for each of your coins._ (one slide summarizing each coin)

5. Coordinate with your team, to combine each of your member's coin descriptions into a single presentation.  Please develop a common format or template for presenting your coin summaries so that the common information for a coin (e.g. TVL) is shown in the same format. (Your slide summaries should look the same!)

6. Be prepared to present your team presentation to your client in class on Thursday 9/16.


##$\color{blue}{Exercise~2} (10 pts)$ 

1. Perform a "creative exploration" of some aspect of the Defi data that you find interesting.   Add your work in this notebook. _Your work should include at least one data visualization._

2. Put your work in this notebook.

3. Write a paragraph describing your findings in the context of DeFi. 

4. Be prepared to share (2 minutes) in team meeting.

5. _You'll receive extra credit for work that goes "above and beyond" this assignment! _

```{r}
stableBorrows <- df %>%
  filter(borrowRateMode == 'Stable') %>%
  select(amount, borrowRate, reserve, reservePriceETH, reservePriceUSD, amountUSD) %>%
  group_by(reserve) %>%
  summarise(meanBorrowRate = mean(borrowRate), meanAmountUSD = mean(amountUSD)) %>%
  arrange(-meanAmountUSD)

kable(stableBorrows)
ggplot(stableBorrows, aes(meanBorrowRate, meanAmountUSD)) + geom_point() + ggtitle("Stable Rate Borrows by Coin") + xlab("Average Borrow Rate (% APY)") + ylab("Average Amount Borrowed (USD)")
```

```{r}
variableBorrows <- df %>%
  filter(borrowRateMode == 'Variable') %>%
  select(amount, borrowRate, reserve, reservePriceETH, reservePriceUSD, amountUSD, variableBorrowRate) %>%
  group_by(reserve) %>%
  summarise(meanBorrowRate = mean(borrowRate), meanAmountUSD = mean(amountUSD)) %>%
  arrange(-meanAmountUSD)
  
kable(variableBorrows)
ggplot(variableBorrows, aes(meanBorrowRate, meanAmountUSD)) + geom_point() + ggtitle("Variable Rate Borrows by Coin") + xlab("Average Borrow Rate (% APY)") + ylab("Average Amount Borrowed (USD)")
```
```{r}
allCoins <- df %>%
  group_by(reserve) %>%
  count(reserve) %>%
  arrange(-n)

kable(allCoins)
```


```{r}
stableCoins <- df %>%
  filter(reserve == 'USDT' | reserve == 'USDC' | reserve == 'GUSD' | reserve == 'DAI' | reserve == 'WBTC' | reserve == 'SUSD' | reserve == 'TUSD') %>%
  group_by(reserve) %>%
  count(reserve) %>%
  arrange(-n)

kable(stableCoins)
```
```{r}
stableCoinTransactions <- df %>%
  filter(reserve %in% stableCoins$reserve) %>%
  group_by(reserve) %>% 
  count(type) %>% 
  mutate(percent = n/sum(n)*100)

kable(stableCoinTransactions)
```
```{r}
ammCoins <- df %>%
  filter(grepl('Amm', reserve))%>%
  group_by(reserve) %>%
  count(reserve) %>%
  arrange(-n)

ammCoinTransactions <- df %>%
  filter(reserve %in% ammCoins$reserve) %>%
  group_by(reserve) %>% 
  count(type) %>% 
  mutate(percent = n/sum(n)*100)
  
kable(ammCoinTransactions)
```


```{r}
unstableCoins <- df%>%
  filter(! reserve %in% stableCoins$reserve & ! reserve %in% ammCoins$reserve) %>%
  group_by(reserve) %>%
  count(reserve) %>%
  arrange(-n)

unstableCoinTransactions <- df %>%
  filter(reserve %in% unstableCoins$reserve) %>%
  group_by(reserve) %>% 
  count(type) %>% 
  mutate(percent = n/sum(n)*100)

kable(unstableCoinTransactions)
```
```{r}
borrowRateSwaps <- df %>%
  filter(type == 'swap') %>%
  select(borrowRateModeFrom, borrowRateModeTo, reserve, variableBorrowRate, stableBorrowRate) %>%
  group_by(borrowRateModeFrom)

variableSwaps <- borrowRateSwaps %>%
  filter(borrowRateModeFrom == 'Variable')

stableSwaps <- borrowRateSwaps %>%
  filter(borrowRateModeFrom  == 'Stable')

ggplot(variableSwaps, aes(variableBorrowRate, stableBorrowRate)) + geom_point() + ggtitle("Swaps from Variable to Stable Borrows")
ggplot(stableSwaps, aes(stableBorrowRate, variableBorrowRate)) + geom_point() + ggtitle("Swaps from Stable to Variable Borrows")

```
